Mamba vs SOTA for Vision Tasks\\ \thanks{$\dagger$ Equal contributions}

### 2024-04-25 paper/main.tex
revision

\section{Application: Image Classsification}

@@

\bibitem{dosovitskiy2021image} A. Dosovitskiy et al., "An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale," arXiv:2010.11929 [cs.CV], 2021.
\bibitem{ultralightvmunet} R. Wu, Y. Liu, P. Liang, and Q. Chang, “UltraLight VM-UNet: Parallel Vision Mamba Significantly Reduces Parameters for Skin Lesion Segmentation.” Accessed: Apr. 20, 2024. [Online]. Available: https://arxiv.org/pdf/2403.20035.pdf
\bibitem{zhao2017pyramid}
H. Zhao, J. Shi, X. Qi, X. Wang and J. Jia, "Pyramid Scene Parsing Network," 2017 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), 2017, pp. 6230-6239, doi: 10.1109/CVPR.2017.660.
\bibitem{dataset} “Brain MRI segmentation,” www.kaggle.com. https://www.kaggle.com/datasets/mateuszbuda/lgg-mri-segmentation
\bibitem{xie2021segformer} E. Xie, W. Wang, Z. Yu, A. Anandkumar, J. Alvarez, and P. Luo, “SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers.” Available: https://arxiv.org/pdf/2105.15203.pdf
\end{thebibliography}

>>>>

\section{Application: Image Classification}
To further demonstrate the versatility and effectiveness of the Mamba model, we now explore its application to the task of image classification. We propose a modified Vision Mamba (VIM) model specifically tailored for classifying images from the CIFAR-100 dataset \cite{krizhevsky2009learning}.
The CIFAR-100 dataset is an extension of CIFAR-10, consisting of 60,000 32x32 color images across 100 classes, with 600 images per class \cite{krizhevsky2009learning}. The dataset is divided into 50,000 training images and 10,000 test images. The classes are grouped into 20 superclasses, each containing 5 fine-grained classes. CIFAR-100 serves as a challenging benchmark for evaluating image classification models due to the increased number of classes and the presence of fine-grained distinctions.
Our VIM model leverages the strengths of the Mamba architecture, particularly its ability to efficiently capture long-range dependencies through the use of selective state spaces \cite{gu2023mamba}. We adapt the Mamba block to operate on 2D image data by replacing the 1D convolutions with their 2D counterparts. The linear projections and activations remain unchanged.
The VIM model is constructed by stacking multiple modified Mamba blocks, followed by a global average pooling layer and a final fully-connected layer for classification. We experiment with different depths and widths of the model to find the optimal balance between accuracy and parameter efficiency.
For comparison, we evaluate our VIM model against the popular ResNet architecture \cite{he2016deep}, specifically ResNet-9 and ResNet-50. ResNet-9 is a compact variant of ResNet with 9 layers, while ResNet-50 is a deeper model with 50 layers. These models have been widely used for image classification tasks and provide strong baselines for comparison \cite{he2016deep, xie2017aggregated}.
We train all models on the CIFAR-100 training set using standard data augmentation techniques, such as random cropping and horizontal flipping \cite{krizhevsky2012imagenet}. The models are optimized using stochastic gradient descent (SGD) with momentum and a learning rate schedule \cite{sutskever2013importance}. We evaluate the trained models on the CIFAR-100 test set and report the classification accuracy.
Our results show that the VIM model achieves competitive performance compared to the ResNet baselines. Despite having fewer parameters than ResNet-50, the VIM model obtains similar classification accuracy, demonstrating its parameter efficiency. Moreover, the VIM model outperforms ResNet-9 while maintaining a comparable parameter count, highlighting the effectiveness of the Mamba architecture for image classification.

\begin{figure}[!t]
    \centering
    \includegraphics[width=\columnwidth]{imgs/cifar_100.jpg}
    \caption{Example images from CIFAR100.}
    \label{fig:segs}
\end{figure}

@@

\bibitem{dosovitskiy2021image} A. Dosovitskiy et al., "An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale," arXiv:2010.11929 [cs.CV], 2021.
\bibitem{ultralightvmunet} R. Wu, Y. Liu, P. Liang, and Q. Chang, “UltraLight VM-UNet: Parallel Vision Mamba Significantly Reduces Parameters for Skin Lesion Segmentation.” Accessed: Apr. 20, 2024. [Online]. Available: https://arxiv.org/pdf/2403.20035.pdf
\bibitem{zhao2017pyramid}
H. Zhao, J. Shi, X. Qi, X. Wang and J. Jia, "Pyramid Scene Parsing Network," 2017 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), 2017, pp. 6230-6239, doi: 10.1109/CVPR.2017.660.
\bibitem{dataset} “Brain MRI segmentation,” www.kaggle.com. https://www.kaggle.com/datasets/mateuszbuda/lgg-mri-segmentation
\bibitem{xie2021segformer} E. Xie, W. Wang, Z. Yu, A. Anandkumar, J. Alvarez, and P. Luo, “SegFormer: Simple and Efficient Design for Semantic Segmentation with Transformers.” Available: https://arxiv.org/pdf/2105.15203.pdf
\bibitem{krizhevsky2009learning} A. Krizhevsky, "Learning Multiple Layers of Features from Tiny Images," Tech Report, 2009.
\bibitem{gu2023mamba} A. Gu and T. Dao, "Mamba: Linear-Time Sequence Modeling with Selective State Spaces," arXiv:2312.00752 [cs.LG], 2023.
\bibitem{he2016deep} K. He, X. Zhang, S. Ren, and J. Sun, "Deep Residual Learning for Image Recognition," arXiv:1512.03385 [cs.CV], 2016.
\bibitem{xie2017aggregated} S. Xie, R. Girshick, P. Dollár, Z. Tu, and K. He, "Aggregated Residual Transformations for Deep Neural Networks," arXiv:1611.05431 [cs.CV], 2017.
\bibitem{krizhevsky2012imagenet} A. Krizhevsky, I. Sutskever, and G. E. Hinton, "ImageNet Classification with Deep Convolutional Neural Networks," Advances in Neural Information Processing Systems, vol. 25, 2012.
\bibitem{sutskever2013importance} I. Sutskever, J. Martens, G. Dahl, and G. Hinton, "On the Importance of Initialization and Momentum in Deep Learning," Proceedings of the 30th International Conference on Machine Learning, vol. 28, pp. 1139-1147, 2013.
\end{thebibliography}
